A Systematic Literature Review on Role-Based Access Control

### 2013-02-03 paper/sections/process.tex
Updates to latest intro and process documentation.

% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{00A0}{\nobreakspace}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{times}
\usepackage[Bjarne]{fncychap}
\usepackage{longtable}
\usepackage{sphinx}
\usepackage{multirow}

\title{RBAC-SLR Documentation}
\date{February 02, 2013}
\release{0.1}
\author{Eric D. Helms}
\newcommand{\sphinxlogo}{}
\renewcommand{\releasename}{Release}
\makeindex

\makeatletter
\def\PYG@reset{\let\PYG@it=\relax \let\PYG@bf=\relax%
    \let\PYG@ul=\relax \let\PYG@tc=\relax%
    \let\PYG@bc=\relax \let\PYG@ff=\relax}
\def\PYG@tok#1{\csname PYG@tok@#1\endcsname}
\def\PYG@toks#1+{\ifx\relax#1\empty\else%
    \PYG@tok{#1}\expandafter\PYG@toks\fi}
\def\PYG@do#1{\PYG@bc{\PYG@tc{\PYG@ul{%
    \PYG@it{\PYG@bf{\PYG@ff{#1}}}}}}}
\def\PYG#1#2{\PYG@reset\PYG@toks#1+\relax+\PYG@do{#2}}

\expandafter\def\csname PYG@tok@gd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PYG@tok@gu\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PYG@tok@gt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.25,0.82}{##1}}}
\expandafter\def\csname PYG@tok@gs\endcsname{\let\PYG@bf=\textbf}
\expandafter\def\csname PYG@tok@gr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PYG@tok@cm\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@vg\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@m\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@mh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@cs\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{1.00,0.94,0.94}{\strut ##1}}}
\expandafter\def\csname PYG@tok@ge\endcsname{\let\PYG@it=\textit}
\expandafter\def\csname PYG@tok@vc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@il\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@go\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.19,0.19,0.19}{##1}}}
\expandafter\def\csname PYG@tok@cp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@gi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PYG@tok@gh\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PYG@tok@ni\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.84,0.33,0.22}{##1}}}
\expandafter\def\csname PYG@tok@nl\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.13,0.44}{##1}}}
\expandafter\def\csname PYG@tok@nn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
\expandafter\def\csname PYG@tok@no\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.38,0.68,0.84}{##1}}}
\expandafter\def\csname PYG@tok@na\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@nb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@nc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
\expandafter\def\csname PYG@tok@nd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.33,0.33,0.33}{##1}}}
\expandafter\def\csname PYG@tok@ne\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@nf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.49}{##1}}}
\expandafter\def\csname PYG@tok@si\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.44,0.63,0.82}{##1}}}
\expandafter\def\csname PYG@tok@s2\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@vi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@nt\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.45}{##1}}}
\expandafter\def\csname PYG@tok@nv\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@s1\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@gp\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
\expandafter\def\csname PYG@tok@sh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@ow\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@sx\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
\expandafter\def\csname PYG@tok@bp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@c1\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@kc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@c\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@mf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@err\endcsname{\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PYG@tok@kd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@ss\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.32,0.47,0.09}{##1}}}
\expandafter\def\csname PYG@tok@sr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.14,0.33,0.53}{##1}}}
\expandafter\def\csname PYG@tok@mo\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@mi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@kn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@o\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PYG@tok@kr\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@s\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@kp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@w\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PYG@tok@kt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.56,0.13,0.00}{##1}}}
\expandafter\def\csname PYG@tok@sc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@sb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@k\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@se\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@sd\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}

\def\PYGZbs{\char`\\}
\def\PYGZus{\char`\_}
\def\PYGZob{\char`\{}
\def\PYGZcb{\char`\}}
\def\PYGZca{\char`\^}
\def\PYGZam{\char`\&}
\def\PYGZlt{\char`\<}
\def\PYGZgt{\char`\>}
\def\PYGZsh{\char`\#}
\def\PYGZpc{\char`\%}
\def\PYGZdl{\char`\$}
\def\PYGZti{\char`\~}
% for compatibility with earlier versions
\def\PYGZat{@}
\def\PYGZlb{[}
\def\PYGZrb{]}
\makeatother

\begin{document}

\maketitle
\tableofcontents
\phantomsection\label{index::doc}

Contents:

\chapter{1. Introduction}
\label{intro:introduction}\label{intro:systematic-literature-review-of-rbac-analysis-notes}\label{intro::doc}
Why is the base model of RBAC extended by newer models?

Role-based access control (RBAC) models {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}} become popularly used to govern access to critical resources.  In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization. RBAC assigns permissions of specific actions on resources to roles instead of individual users.  Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.

RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.  RBAC has advantages in maintaining and managing organization's security policies.  For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.

RBAC is first introduced in 1990s, NIST proposed standard RBAC model {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}}
Standard RBAC model considers only role-user association and role hierarchy.
Since standard RBAC model has limitations such as specifying environmental constraints or context information
Researchers developed extended models of RBAC to overcome the limitations.
However, as researchers often develop their own specialized extended models of RBAC,
their research cannot be generalized or compared with other research work appropriately.
As a result, researchers could take time on reinventing the wheel.
But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?

The goal of this work is to synthesize available research results on extended models of RBAC. We analyze their extended features and claimed research contributions to find limitations of current RBAC models and what extent of extended features by
comparing with similar research work.
We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
\begin{description}
\item[{Our research give benefits to a community as follows:}] \leavevmode\begin{itemize}
\item {} 
Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.

\item {} 
Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.

\item {} 
Our work shows a criteria in comparison among research results.

\item {} 
Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.

\end{itemize}

\end{description}

\chapter{2. Methodology and Process}
\label{process:methodology-and-process}\label{process::doc}
For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
\begin{itemize}
\item {} 
Google Scholar - \href{http://scholar.google.com}{http://scholar.google.com}

\item {} 
IEEExplore     - \href{http://ieeexplore.ieee.org}{http://ieeexplore.ieee.org}

\item {} 
ACM Portal     - \href{http://dl.acm.org}{http://dl.acm.org}

\item {} 
CiteSeerX      - \href{http://http://citeseerx.ist.psu.edu/index}{http://http://citeseerx.ist.psu.edu/index}

\end{itemize}

For each of the criteria below, a search was performed on each of the search engines for a total of 12 data sets.  The criteria used were:
\begin{itemize}
\item {} 
role based access control

\item {} 
RBAC

\item {} 
role-based access control

\end{itemize}

The search performed was done in an automated way using a set of scripts to query and collect data from each search engine with the criteria string as input.  For each criteria for each search engine, the results were captured until a stopping criteria was met.  Each run was performed as follows:
\begin{enumerate}
\item {} 
Remote call to search engine with current search start position and the current search criteria.

\item {} 
Parse results and extract paper title, authors and year of publication.

\item {} 
Compare results against stopping criteria.

\end{enumerate}
\begin{itemize}
\item {} 
If stopping criteria met, stop search.

\item {} 
If stopping criteria not met, increase search position by number of results and return to step 1.

\end{itemize}

@@

There were a total of three elimination rounds.  The first round was based solely on title, the second on reading of the abstract and the last round was based on a full read of the paper and comparison to the research questions outlined.  Each round was performed as follows:
\begin{itemize}
\item {} 
Each reviewer independently classified papers as relevant, irrelevant or uncertain.

\item {} 
Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.

\item {} 
Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.  From this discussion, papers were either thrown out or kept until the next round of the review.  Ties were broken by an indepedent party.

\end{itemize}

@@

Google Scholar
 & 
651
 & 
213
 & 
435
 & 
1299
\\\hline

ACM Portal
 & 
500
 & 
20
 & 
720
 & 
1240
\\\hline

IEEExplore
 & 
200
 & 
40
 & 
230
 & 
470
\\\hline

CiteSeerX
 & 
100
 & 
100
 & 
150
 & 
350
\\\hline
 &  &  &  & \\\hline

Totals
 & 
1451
 & 
373
 & 
1535
 & 
3359
\\\hline

@@

\section{Research Questions}
\label{process:research-questions}
Based on our intial intent for the systematic literature review, and notes taken during the third phase of elimination, we have identified the following research questions:
\begin{itemize}
\item {} 
RQ1. What are the deficiencies in current RBAC model to propose extended RBAC models?

\item {} 
RQ2. What are the motivations behind RBAC extensions?

\item {} 
RQ3. What are the categorizations of RBAC model proposed?

\item {} 
RQ4. For each categorization, what are the extended features of RBAC model proposed?

\item {} 
RQ5. Do these models have corresponding implementations in practice?

\item {} 
RQ6. How are extended RBAC models evaluated in theory and in practice?

\item {} 
RQ7. Are there any commonalties or generalizations across all categorizations?

\item {} 
RQ8. What domains or scenarios serve as inspiration for these extensions?

\end{itemize}

\chapter{RQ1: What are the deficiencies in current RBAC model?}
\label{research_questions/RQ_1:rq1-what-are-the-deficiencies-in-current-rbac-model}\label{research_questions/RQ_1::doc}

\section{Common Themes}
\label{research_questions/RQ_1:common-themes}\begin{itemize}
\item {} 
\end{itemize}

\section{By Category}
\label{research_questions/RQ_1:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
no support for partial inheritance

\item {} 
inability to support permission assignment constraints

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
Missing ability to apply constraints between role and permissions and enforce

\item {} 
Missing context that defines if a role can be activated at execution time

\item {} 
Role activation on a per object or process execution rather than session based

\item {} 
No allowance for context to check permission at execution moment based on context

\item {} 
activation of roles based on current context rather than session intiation

\item {} 
conflict resolution

\item {} 
RBAC cannot handle dynamic authorization and entities

\item {} 
Role cannot be resource dependent

\item {} 
RBAC is too simple

\item {} 
Missing granularity

\item {} 
Inability to handle dynamic nature of multiple projects with multiple resources

\item {} 
Too simple

\item {} 
Lacks context

\item {} 
Handling authorization between applications and within can be different

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
RBAC cannot handle cross-organizational roles and access in an administratively easy way

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
Privacy regulations are complex, with many entities, RBAC lacks ability to handle all of these

\item {} 
Context, conflict detection and resolution

\item {} 
RBAC lacks ability to define purpose

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
No support for multiple security domains

\item {} 
No support for same security, different applications

\item {} 
Lack of flexibility

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
Lack of contextual information

\item {} 
Role needs a defined schema

\item {} 
Role activation based on context

\item {} 
Giving a permission an operation

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
Needs an entity to represent a task and connect a role and a permission through it

\item {} 
Need for team and tasks

\item {} 
Need for entity between permission and roles to hold constraint and allow for dynamic situations and entities

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
RBAC can't handle time based constraints for role enablement

\end{itemize}

\end{description}

\chapter{RQ2: What are the motivations behind RBAC extensions?}
\label{research_questions/RQ_2:rq2-what-are-the-motivations-behind-rbac-extensions}\label{research_questions/RQ_2::doc}

\section{Common Themes}
\label{research_questions/RQ_2:common-themes}\begin{itemize}
\item {} 
Roles, users and permissions need context associations

\item {} 
Location and time constraints as well as contexts needed

\item {} 
RBAC lacks ability to link in objects

\item {} 
RBAC does not handle single user base and multiple organizations/applications

\item {} 
Role enablement based on context

\item {} 
User granting/revoking of roles based on context

\end{itemize}

\section{By Category}
\label{research_questions/RQ_2:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
No support for partial inheritance

\item {} 
Permission assignment constraints

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
Relationship between object and viewer can't be represented

\item {} 
Object content matters

\item {} 
Location and time constraints on the user

\item {} 
Departmental context around access

\item {} 
RBAC does not handle many small services to wrap around them

\item {} 
RBAC lacks security principle enforcement

\item {} 
Lack of specifying context, and permission checks of dynamic evaluation time rules

\item {} 
RBAC is static and lacks ability to handle dynamic authorization and definitions

\item {} 
RBAC ignores context

\item {} 
Role granulatiry not well controlled

\item {} 
Roles should depend on the resources

\item {} 
RBAC cannot handle multiple projects with multiple tasks and differing roles in each system but the same user

\item {} 
Lack of context support

\item {} 
Too simple for dynamic environments

\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
Support cross-domain delegation

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
Reduce administrative complexity of RBAC for multi-organizational

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
Lack of components, purpose binding, conditions, obligations to handle privacy needs

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
Need to support multiple securtiy domains

\item {} 
Need to support same security with different applications

\item {} 
Role heirarchy adds complexity and reduces flexibility

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
Lack of contextual information with roles, users and permissions

\item {} 
Role needs a schema to allow specification of attributes

\item {} 
Location based security scenarios

\item {} 
Role activation based on location

\item {} 
Support physical and logical locations

\item {} 
Attaching operations to permissions

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
RBAC doesn't inherently support location and time constraints

\item {} 
Spatial and temporal awareness of roles, inheritance and permissions

\item {} 
Context mechanisms needed

\item {} 
Physical and logical locations, hybrids

\item {} 
Role assignment based off time and space, as well as permissions

\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
RBAC can't handle large enterprises with multiple applications and single user base

\item {} 
Security requirements for enterprises (look up these)

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
Workflow based organizations need time base constraint for role enablement

\end{itemize}

\end{description}

\chapter{RQ3: What are the categorizations of proposed RBAC extensions?}
\label{research_questions/RQ_3:rq3-what-are-the-categorizations-of-proposed-rbac-extensions}\label{research_questions/RQ_3::doc}
Across the papers found, a number of higher level categories emerged that each proposed model fell in to.  Some of these categories had singular primary papers, while others had multiple extension models proposed that were deemed independent of one another.

For each category, the breakdown of the papers for each category is:

\begin{tabulary}{\linewidth}{|L|L|}
\hline
\textbf{
Paper
} & \textbf{
Category
}\\\hline

Alam, M. and Hafner, M. and Breu, R. 2006
 & 
Constraint
\\\hline

Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001
 & 
Context
\\\hline

Haibo, SHEN and Fan, HONG 2005
 & 
Context
\\\hline

Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000
 & 
Context
\\\hline

Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006
 & 
Context
\\\hline

Motta, G.H.M.B. and Furuie, S.S. 2003
 & 
Context
\\\hline

Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008
 & 
Context
\\\hline

Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004
 & 
Context
\\\hline

Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008
 & 
Context
\\\hline

Thein, N. and others 2011
 & 
Context
\\\hline

Zou, D. and He, L. and Jin, H. and Chen, X. 2009
 & 
Context
\\\hline

Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010
 & 
Delegation
\\\hline

Zhang, Z. and Zhang, X. and Sandhu, R. 2006
 & 
Organizational
\\\hline

Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007
 & 
Privacy
\\\hline

Masoumzadeh, A. and Joshi, J. 2008
 & 
Privacy
\\\hline

Zhao, Y. and Zhao, Y. and Lu, H. 2008
 & 
Resource
\\\hline

Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005
 & 
Spatial
\\\hline

Ray, I. and Kumar, M. and Yu, L. 2006
 & 
Spatial
\\\hline

Hansen, F. and Oleshchuk, V. 2003
 & 
Spatial
\\\hline

Aich, S. and Sural, S. and Majumdar, A. 2007
 & 
Spatio-Temporal
\\\hline

Chen, L. and Crampton, J. 2008
 & 
Spatio-Temporal
\\\hline

Samuel, A. and Ghafoor, A. and Bertino, E. 2007
 & 
Spatio-Temporal
\\\hline

Chandran, S. and Joshi, J. 2005
 & 
Spatio-Temporal
\\\hline

Ray, I. and Toahchoodee, M. 2007
 & 
Spatio-Temporal
\\\hline

Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009
 & 
Spatio-Temporal
\\\hline

Yao, L. and Kong, X. and Xu, Z. 2008
 & 
Task
\\\hline

ZHANG, S. and CHEN, X. and HOU, G. 2009
 & 
Task
\\\hline

Oh, S. and Park, S. 2003
 & 
Task
\\\hline

Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005
 & 
Temporal
\\\hline
\end{tabulary}

\section{By Category}
\label{research_questions/RQ_3:by-category}\begin{description}
\item[{Constraint (1)}] \leavevmode
a limitation or restriction

\item[{Context (8)}] \leavevmode
circumstances in which an event occurs

\item[{Delegation (1)}] \leavevmode
the act of empowering to act for another

\item[{Organizational (1)}] \leavevmode
of or relating to an organization

\item[{Privacy (1)}] \leavevmode
freedom from unauthorized intrusion

\item[{Resource (1)}] \leavevmode
an available supply than be drawn on when needed

\item[{Spatial (3)}] \leavevmode
of or relating to facility in percieving relations in space

\item[{Spatio/Temporal (6)}] \leavevmode
combining space and time

\item[{Task (3)}] \leavevmode
a usually assigned piece of work often to be finished within a certain time

\item[{Temporal (1)}] \leavevmode
of or relating to the sequence of time or to a particular time

\end{description}

{\hfill\scalebox{1.000000}{\includegraphics{categories_map.png}}\hfill}

\chapter{RQ4: For each categorization, what are the extended features of RBAC proposed?}
\label{research_questions/RQ_4::doc}\label{research_questions/RQ_4:rq4-for-each-categorization-what-are-the-extended-features-of-rbac-proposed}
Is this the same as RQ\_1 ?

\section{Common Themes}
\label{research_questions/RQ_4:common-themes}\begin{itemize}
\item {} 
\end{itemize}

\section{By Category}
\label{research_questions/RQ_4:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\end{description}

\chapter{RQ5. Do these models have corresponding implementations in practice?}
\label{research_questions/RQ_5::doc}\label{research_questions/RQ_5:rq5-do-these-models-have-corresponding-implementations-in-practice}
When designing and proposing a model targeted at a feature that is rooted in practical
usage by real software systems, bringing the model to life is strong evidence that the
proposed model can work in practice.  The concept of authorization, and access control
is rooted in a business need.  Thus, any access control model needs to be feasible
in the real world not just on paper.  We analyzed the primary sources to see how many
proposed models actually had implmenetations associated with them.  And quantified the
type of implementation.  Whether the implementation was for a real system, for a prototype
and/or used in a production environment.

\section{Common Themes}
\label{research_questions/RQ_5:common-themes}\begin{itemize}
\item {} 
21 No's

\item {} 
8 Yes's

\item {} 
Half the implementation's were prototypes

\end{itemize}

\section{By Category}
\label{research_questions/RQ_5:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
No - system architecture provided

\item {} 
No

\item {} 
Yes - prototype with VB6 and MS Access97

\item {} 
Yes - prototype system, system architecture, algorithms implemented, simulation scenarios

\item {} 
Yes - on an EHR system

\item {} 
Yes - prototype using XML container

\item {} 
Yes - web based API agent

\item {} 
No

\item {} 
No

\item {} 
No

\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\item {} 
No

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
Yes - integrated with telemedicine application, web based

\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
Yes - mentions of systems that exist using the model

\item {} 
Yes - prototype system

\item {} 
No

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\end{description}

\chapter{RQ6. How are extended RBAC models evaluated in theory and in practice?}
\label{research_questions/RQ_6:rq6-how-are-extended-rbac-models-evaluated-in-theory-and-in-practice}\label{research_questions/RQ_6::doc}

\section{Common Themes}
\label{research_questions/RQ_6:common-themes}\begin{itemize}
\item {} 
12 had no evaluation whatsoever

\item {} 
8 present example scenarios and how their model would apply

\item {} 
6 presented performance and time complexity analysis

\end{itemize}

\section{By Category}
\label{research_questions/RQ_6:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
No evaluation

\end{itemize}

>>>>

\section{Methodology and Process} \label{sec:process}

For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
\begin{itemize}
\item {} 
Google Scholar - \href{http://scholar.google.com}{http://scholar.google.com}
\item {} 
IEEExplore     - \href{http://ieeexplore.ieee.org}{http://ieeexplore.ieee.org}
\item {} 
ACM Portal     - \href{http://dl.acm.org}{http://dl.acm.org}
\item {} 
CiteSeerX      - \href{http://http://citeseerx.ist.psu.edu/index}{http://http://citeseerx.ist.psu.edu/index}
\end{itemize}

For each of the criteria below, a search was performed on each of the search engines for a total of 12 data sets.  The criteria used were:
\begin{itemize}
\item {} 
role based access control
\item {} 
RBAC
\item {} 
role-based access control
\end{itemize}

The search performed was done in an automated way using a set of scripts to query and collect data from each search engine with the criteria string as input.  For each criteria for each search engine, the results were captured until a stopping criteria was met.  Each run was performed as follows:
\begin{enumerate}
\item {} 
Remote call to search engine with current search start position and the current search criteria.
\item {} 
Parse results and extract paper title, authors and year of publication.
\item {} 
Compare results against stopping criteria.
\end{enumerate}

\begin{itemize}
\item {} 
If stopping criteria met, stop search.
\item {} 
If stopping criteria not met, increase search position by number of results and return to step 1.
\end{itemize}

@@

There were a total of three elimination rounds.  The first round was based solely on title, the second on reading of the abstract and the last round was based on a full read of the paper and comparison to the research questions outlined.  Each round was performed as follows:

\begin{itemize}
\item {} 
Each reviewer independently classified papers as relevant, irrelevant or uncertain.
\item {} 
Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
\item {} 
Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.  From this discussion, papers were either thrown out or kept until the next round of the review.  Ties were broken by an indepedent party.
\end{itemize}

@@

Google Scholar & 651 & 213 & 435 & 1299
\\\hline
ACM Portal & 500 & 20 & 720 & 1240
\\\hline
IEEExplore & 200 & 40 & 230 & 470
\\\hline
CiteSeerX & 100 & 100 & 150 & 350
\\\hline
 &  &  &  & 
\\\hline
Totals & 1451 & 373 & 1535 & 3359
\\\hline

@@
